---
title: "Barchart futures data"
output: html_document
date: "2023-07-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

Show me the money. Time to wrangle Barchart futures data. Let's see if I can make practice out of theory.

I got a free trial of BarChart and have downloaded some data. Reminders on the month codes:

Check these are good, as I hacked off the last row with some "watermarking" from barchart
```{r import the futures price data }
zcu23_daily_historical_data_07_08_2023 <- read_csv("zcu23_daily_historical-data-07-08-2023.csv", 
     col_types = cols(Time = col_date(format = "%m/%d/%Y")))
zch24_daily_historical_data_07_08_2023 <- read_csv("zch24_daily_historical-data-07-08-2023.csv", 
                                                   col_types = cols(Time = col_date(format = "%m/%d/%Y")))

```

```{r plot of daily ZCH future price}
 ggplot(data = zch24_daily_historical_data_07_08_2023, aes(x = Time, y = High, color = "High"))+geom_line()+theme_minimal()+geom_line(aes(x=Time, y= Open, color = "Open"))+geom_line(aes(x=Time, y= Last, color = "Last"))
```

Can compare the two series
```{r plot of ZCU and ZCH}
ggplot(data = zch24_daily_historical_data_07_08_2023, aes(x = Time, y = High, color = "24"))+geom_line()+theme_minimal()+geom_line(data = zcu23_daily_historical_data_07_08_2023, aes(x=Time, y = High, color = "23"))
```

Ok, great, as a test case we're in business. I will download a bunch more stuff. Next to-dos: 

- look at the difference between actual corn prices and the next expiring future
- do some options pricing & look at some trades
- move on to value at risk
- make more coffee, it was weak this morning