---
title: "Supply and Demand Balance Sheets"
output: html_document
date: "2023-07-07"
---

Alright. What can I do in R for supply and demand balance sheets? I want to use WASDE's data for the baseline. 

It's easy to get the text data into R, but the format's a bit rough. read_table from reader is good for whitespace-deliminated information. The difficulty here is that WASDE's text has a lot of text that shouldn't be parsed as whitespace delimited. So, what's easier, parsing this txt file, or using the xls file and pulling the tab I want? 

```{r read in txt from wasde download}
library(readr)
wasde0623 <- read_table("wasde0623.txt")
View(wasde0623)
```

I am particularly interested in corn and sugar. For the xls files, "Page 16" is sugar and "Page 12" has corn.

```{r read in xls file from wasde download}
library(readxl)
wasde_sugar_0623 <- read_excel("wasde0623.xls", sheet = "Page 16", skip = 8)

```
Ok, have the ugly file. Do people use xml for a better workflow? this is so ugly.

Well, I can pull out hte stocks to use ratios. 

Should I pull in this old historical csv? That might be better.

```{r pull in the historical csv}
hist_wasde <- read_csv("oce-wasde-report-data-2023-06.csv")
```
This is refreshing. Formatting! wow! 

Let's select sugar and plot some stuff. Unfortunately, this "historical" data goes all the way back to (gasp) 2021. Need to pull in some older files from the website.

```{r 2020s sugar info}
hist_2020s_sugar = hist_wasde[hist_wasde$Commodity == "Sugar",]
oce_wasde_report_data_2016_01_to_2020_12 <- read_csv("oce-wasde-report-data-2016-01-to-2020-12.csv")
oce_wasde_report_data_2010_04_to_2015_12 <- read_csv("oce-wasde-report-data-2010-04-to-2015-12.csv")
hist_2016p_sugar = oce_wasde_report_data_2016_01_to_2020_12[oce_wasde_report_data_2016_01_to_2020_12$Commodity == "Sugar",]
hist_2010p_sugar = oce_wasde_report_data_2010_04_to_2015_12[oce_wasde_report_data_2010_04_to_2015_12$Commodity == "Sugar",]
```
Alrighty then. There's a lot of balance sheet info, but maybe we can be really lazy and pull "Stocks to Use Ratio". I want to plot this over time. Maybe color by ProjEstFlag? It's all annual... How do I want to arrange the dates? These reports are frequent but projecting forward for each market year. 

```{r a crappy plot of estimates and projections}
sugar_2010p_stu = hist_2010p_sugar[hist_2010p_sugar$Attribute == "Stocks to Use Ratio",]
ggplot(data = sugar_2010p_stu[sugar_2010p_stu$Region == "United States",], aes(x = MarketYear, y = Value, color = ProjEstFlag))+geom_point()+theme_minimal()
```
This is evocative, in a weird way. I don't like how it's showing up, though. Hm. 

Continue this EDA

```{r historical beginning stocks sugar}
ggplot(data = hist_2010p_sugar %>% filter(Region == "United States", Attribute == "Beginning Stocks", is.na(ProjEstFlag)), aes(x = MarketYear, y = Value)) + geom_point() + geom_point(data = hist_2016p_sugar %>% filter(Region == "United States", Attribute == "Beginning Stocks", is.na(ProjEstFlag)), aes(x = MarketYear, y = Value))+theme_minimal()
```

```{r historical ending stocks sugar}
ggplot(data = hist_2010p_sugar %>% filter(Region == "United States", Attribute == "Ending Stocks", is.na(ProjEstFlag)), aes(x = MarketYear, y = Value)) + geom_point() + geom_point(data = hist_2016p_sugar %>% filter(Region == "United States", Attribute == "Ending Stocks", is.na(ProjEstFlag)), aes(x = MarketYear, y = Value))+theme_minimal()
```


```{r stocks to use historical}

ggplot(data = hist_2010p_sugar %>% filter(Region == "United States", Attribute == "Stocks to Use Ratio", is.na(ProjEstFlag)), aes(x = MarketYear, y = Value)) + geom_point() + geom_point(data = hist_2016p_sugar %>% filter(Region == "United States", Attribute == "Stocks to Use Ratio", is.na(ProjEstFlag)), aes(x = MarketYear, y = Value))+theme_minimal() 

```

Ok, fine, that's all nice. Show me the money. Where's the money? and also, this is only exploring US numbers. Mexico numbers have slightly different format, I played with that as well. 
