---
title: "Value at Risk R version"
output: html_document
date: "2023-07-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("quantmod")
library("tidyverse")
```
As a first step, let's work on a tutorial for using quantmod to compute monte carlo VaR. The tutorial is at https://towardsdev.com/monte-carlo-value-at-risk-var-in-r-960e62f0f53f, there is little original in this file. 
```{r tutorial on var via monte carlo }
tsla <- quantmod::getSymbols("TSLA", from = base::as.Date("2021-12-01"), to = base::as.Date("2023-07-09"), auto.assign = F)

## Calculate percent change on TSLA close prices
tsla <- base::data.frame(Date = zoo::index(tsla), zoo::coredata(tsla)) %>% # Convert to a data frame because dplyr::mutate() cannot be applies to an 'xts' object and keep index as date column
  dplyr::arrange(dplyr::desc(Date)) %>% # Sort dates by descending order to ensure we calculate percent change correctly
  dplyr::mutate(Pct_change = (TSLA.Close / dplyr::lead(TSLA.Close)) - 1) # Calculate percent change
```

This gives us our dataframe with percent change. Now let's do some housekeeping:

```{r boilerplate}

## Calculate number of periods
N <- base::nrow(tsla)

## Calculate price volatility
vol <- stats::sd(stats::na.omit(tsla$Pct_change)) * sqrt(252)
```


Run the simulations. The price of Tesla stock has changed quite a bit since the author wrote the tutorial, sigh. 

```{r simulations}

## Run simulations
simulations <- base::list() # Initialize list to hold the simulation output
for (i in 1:300) { # Determine number of simulations to run
  tmp <- base::rep(278, N) # replacing  with a smaller number #956.735, N) # Initialize price with realtime quote (from original VaR post)
  for (t in 2:N) {
    tmp[t] <- tmp[t - 1] * (1 + (((stats::runif(1) - 0.5) * vol) / N)) # Calculate forward price
  }
## Note that this in itself is a choice on how to do the random walks here
  ## Probaly not appropriate for what I want
  simulations <- base::append(simulations, base::list(base::data.frame("Run" = base::rep(i, N), "DaysForward" = 1:N, "Price" = tmp)))
}
simulations <- dplyr::bind_rows(simulations) # Bind back into a data frame

## Plot the simulations
ggplot2::ggplot(simulations, ggplot2::aes(x = DaysForward, y = Price, color = base::as.factor(Run))) + ggplot2::geom_line() +theme_minimal()+ theme(legend.position="none")
```




```{r get end price}

## Select final price of all runs
last_price <- simulations %>% dplyr::filter(DaysForward == N)

```

Now just histogram it and plot/sort/figure out the moeny

```{r histograms}

hist(last_price$Price, main = "TSLA final MonteCarlo price", breaks = 50)
q = quantile(last_price$Price, probs = 0.05)
abline(v=q, col = "red")
```